---
title: "ggplot_HandsOn"
author: "Hong"
date: "2023-02-27"
output: html_document
---

## Setting up my environment:
Notes: setting up my R environment by loading the 'tidyverse' and 'palmer penguins' packages.

```{r loading packages, warning=FALSE}
library(ggplot2)
library(palmerpenguins)
```

## Visualizations
Here we will go through a series of visualizations

### The relationship between body mass and flipper length
The plot shows a positive relationship between the two variables in the three penguin species. 
In other words, the larger the penguin, the longer the flipper. 
As the values on the x-axis increase, the values on the y-axis increase.
Limitation: we can't tell which data points refer to each of three penguin species.

```{r warning=FALSE}
ggplot(data = penguins) + 
  geom_point(mapping = aes(x = flipper_length_mm, y = body_mass_g))

```

### Map the new variable "species" to the new aesthetic "color"
By this way, we can solve the above limitation.

```{r warning=FALSE}
ggplot(data = penguins)+
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g,color=species))
```

### Use shape to highlight the different penguin species
Let's map the variable "species" to the aesthetic "shape".

```{r warning=FALSE}
ggplot(data = penguins)+
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g,shape=species))
```

### Map more than one aesthetic to the same variable
Let's map both color and shape to species

```{r warning=FALSE}
ggplot(data = penguins)+
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g,shape=species, color=species))
```

### Map the alpha aesthetic to species
This is an option for those who like black and white plot. 
This will make some points more transparent or see-through than others. 
This gives us another way to represent each penguin species.

```{r warning=FALSE}
ggplot(data = penguins)+
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g,alpha=species))
```

### Map the aesthetic color to the variable species

```{r warning=FALSE}
ggplot(data = penguins)+
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g,shape=species, color="purple"))
```

### Write code outside of the aes function
Purpose: change the appearance of our overall plot without regard to specific variables.

```{r warning=FALSE}
ggplot(data = penguins)+
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g,shape=species), color="purple")
```

We have the same data but with different visual appearance:

```{r message=FALSE, warning=FALSE}
ggplot(data = penguins)+
  geom_smooth(mapping = aes(x=flipper_length_mm, y=body_mass_g))
```

### Use 2 geoms in the same plot

```{r message=FALSE, warning=FALSE}
ggplot(data = penguins)+
  geom_smooth(mapping = aes(x=flipper_length_mm, y=body_mass_g)) +
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g))
```

### Geom_smooth
We can add the line type aesthetic and map it to the variable species.
Geom_smooth will draw a different line with a different line type for each species.
The plot shows the trend for each species of penguins.

```{r message=FALSE, warning=FALSE}
ggplot(data = penguins)+
  geom_smooth(mapping = aes(x=flipper_length_mm, y=body_mass_g, linetype=species))
```

### Geom_jitter 
This function creates a scatter plot & adds a small amount of random noise to each point of the plot.
Jittering helps us deal with over plotting.

```{r message=FALSE, warning=FALSE}
ggplot(data = penguins)+
  geom_jitter(mapping = aes(x=flipper_length_mm, y=body_mass_g))
```

### Geom_bar 
Use diamonds dataset.
Geom_bar function: R automatically counts how many times each x-value appears in the data,
then shows the counts on the y-axis. The defaults for geom_bar is to count rows.

```{r message=FALSE, warning=FALSE}
ggplot(data=diamonds)+
  geom_bar(mapping = aes(x=cut))
```

### Add color aesthetic to the plot

```{r message=FALSE, warning=FALSE}
ggplot(data=diamonds)+
  geom_bar(mapping = aes(x=cut, color=cut))
```

### Add color to the inside of the bar

```{r message=FALSE, warning=FALSE}
ggplot(data=diamonds)+
  geom_bar(mapping = aes(x=cut, fill=cut))
```

### Stacked bar chart

```{r message=FALSE, warning=FALSE}
ggplot(data=diamonds)+
  geom_bar(mapping = aes(x=cut, fill=clarity))
```

### Facet_wrap 
This function helps us create separate plot for each species.
To add a new layer to the plot, we add a "+" to our code.
Inside the parentheses, type the tilde symbol "~" followed by the variable.

```{r message=FALSE, warning=FALSE}
ggplot(data = penguins)+
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g,color=species))+
  facet_wrap(~species)

```


```{r message=FALSE, warning=FALSE}
ggplot(data = penguins, aes(x=flipper_length_mm,y=body_mass_g))+
  geom_point(aes(color=species))+
  facet_wrap(~species)
```

```{r message=FALSE, warning=FALSE}
ggplot(data = diamonds)+
  geom_bar(mapping = aes(x=color,fill=cut))+
  facet_wrap(~cut)
```

### Facet_grid 
Use this function to facet the plot with 2 variables (sex & species): 
it will split the plot into facets vertically by the values of the 1st variable 
and horizontally by the values of the 2nd variable.

```{r message=FALSE, warning=FALSE}
ggplot(data = penguins)+
  geom_point(mapping = aes(x=flipper_length_mm,y=body_mass_g,color=species))+
  facet_grid(sex~species) 
```

## Annotate the viz

### Add a title to the plot 
To show the relationship between the body mass & flipper length

```{r message=FALSE, warning=FALSE}
ggplot(data = penguins)+
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g,color=species))+
  labs(title="Palmer Penguins: Body Mass vs. Flipper Length")
```

### Add a subtitle to the plot 
To highlight important info

```{r message=FALSE, warning=FALSE}
ggplot(data = penguins)+
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g,color=species))+
  labs(title="Palmer Penguins: Body Mass vs. Flipper Length", subtitle="Sample of Three Penguin Species")
```

### Add a caption
with the same way

```{r message=FALSE, warning=FALSE}
ggplot(data = penguins)+
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g,color=species))+
  labs(title="Palmer Penguins: Body Mass vs. Flipper Length", subtitle="Sample of Three Penguin Species", caption="Data collected by Dr. Kristen Gorman")
```

### annotate function
Use this function to add some text next to the data points

```{r message=FALSE, warning=FALSE}
ggplot(data = penguins)+
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g,color=species))+
  labs(title="Palmer Penguins: Body Mass vs. Flipper Length", subtitle="Sample of Three Penguin Species", 
       caption="Data collected by Dr. Kristen Gorman")+
  annotate("text", x=220, y=3500, label="The Gentoos are the largest")
```

### Add color equals 
followed by the name of the colors:

```{r message=FALSE, warning=FALSE}
ggplot(data = penguins)+
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g,color=species))+
  labs(title="Palmer Penguins: Body Mass vs. Flipper Length", subtitle="Sample of Three Penguin Species", 
       caption="Data collected by Dr. Kristen Gorman")+
  annotate("text", x=220, y=3500, label="The Gentoos are the largest", color="purple")
```

### Edit the font, size, angle...

```{r message=FALSE, warning=FALSE}
ggplot(data = penguins)+
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g,color=species))+
  labs(title="Palmer Penguins: Body Mass vs. Flipper Length", subtitle="Sample of Three Penguin Species", 
       caption="Data collected by Dr. Kristen Gorman")+
  annotate("text", x=220, y=3500, label="The Gentoos are the largest", color="purple", fontface="bold", size=4.5, angle=25)
```

### Other method:
Instead of writing all the codes again, we just call p and add annotation to it.

```{r message=FALSE, warning=FALSE}
p <- ggplot(data = penguins)+
  geom_point(mapping = aes(x=flipper_length_mm, y=body_mass_g,color=species))+
  labs(title="Palmer Penguins: Body Mass vs. Flipper Length", subtitle="Sample of Three Penguin Species", 
       caption="Data collected by Dr. Kristen Gorman")

p+annotate("text",x=220,y=3500,label="The Gentoos are the largest")
```




